Manifold Clustering for Latent Structure Block Models

### 2022-03-08 draft/draft.Rmd
fixed a few bugs

```{r, cache = TRUE}
maxit <- 50
animate <- FALSE
animation.dir <- '~/dev/manifold-block-models/draft'
random.starts <- 2
eps <- 1e-3
out <- lapply(seq(random.starts), function(i) {
  intercept <- (i / random.starts > .5 + eps)
  set.seed(i)
  manifold.clustering(Xhat, 2, 
                      parallel = TRUE, 
                      intercept = intercept, 
                      maxit = maxit,
                      eps = 1e-3,
                      verbose = FALSE,
                      animate = animate,
                      animation.dir = animation.dir,
                      animation.title = i)
})
names(out) <- paste('random', seq(random.starts))
names(out) <- sapply(seq(random.starts), function(i) {
  if (i / random.starts < .5 + eps) {
    paste('random', i, '\n(w/o intercept)')
  } else {
    paste('random', i, '\n(w/ intercept)')
  }
})

out[['spectral clust\n(w/o intercept)']] <- manifold.clustering(
  Xhat, 2, 
  parallel = TRUE, 
  intercept = FALSE, 
  maxit = maxit, 
  eps = 1e-3,
  initialization = 'spectral',
  verbose = FALSE,
  animate = animate,
  animation.dir = animation.dir,
  animation.title = 'spec-clust-no-intercept')

out[['spectral clust\n(w/ intercept)']] <- manifold.clustering(
  Xhat, 2,
  parallel = TRUE,
  intercept = TRUE,
  maxit = maxit,
  eps = 1e-3,
  initialization = 'spectral',
  verbose = FALSE,
  animate = animate,
  animation.dir = animation.dir,
  animation.title = 'spec-clust-w-intercept')

out[['ground truth\n(w/o intercept)']] <- manifold.clustering(
  Xhat, 2,
  parallel = TRUE,
  intercept = FALSE,
  maxit = maxit,
  eps = 1e-3,
  initialization = z,
  verbose = FALSE,
  animate = animate,
  animation.dir = animation.dir,
  animation.title = 'groundtruth-no-intercept')

@@

```{r cache = TRUE}
# set.seed(123)
z.spectral <- spectral.clustering(A)
macaque.out <- manifold.clustering(Xhat, 2, 
                                   # initialization = 'random',
                                   # initialization = z.spectral, 
                                   initialization = z,
                                   intercept = FALSE, 
                                   parallel = TRUE, 
                                   verbose = FALSE, 
                                   animate = animate, 
                                   curve.init = 'x',
                                   animation.title = 'macaque')
```

@@

@@

```{r cache = TRUE}
z.init <- hclust(dist(Xhat), method = 'single') %>% 
  cutree(2)
nonintersect.out <- manifold.clustering(Xhat, 2, 
                                        initialization = z.init,
                                        intercept = TRUE, 
                                        animate = animate,
                                        animation.title = 'nonintersect')
```

@@

ggplot() + 
  geom_point(aes(x = Xhat[, 1], y = Xhat[, 2], 
                 colour = factor(nonintersect.out$z)),
             size = .5) + 
  labs(x = expression(x[1]), y = expression(x[2]), colour = NULL) + 
  theme_bw() + 
  coord_fixed() +
  theme(legend.position = 'none') +
  geom_path(aes(x = Y1[, 1], y = Y1[, 2])) +
  geom_path(aes(x = Y2[, 1], y = Y2[, 2])) + 
  xlim(min(Xhat[, 1]), max(Xhat[, 1])) + ylim(min(Xhat[, 2]), max(Xhat[, 2]))
```

>>>>

```{r}
animate <- FALSE
```

```{r, cache = TRUE}
maxit <- 50
animation.dir <- '~/dev/manifold-block-models/draft'
random.starts <- 2
eps <- 1e-6
normalize <- TRUE

out <- lapply(seq(random.starts), function(i) {
  intercept <- (i / random.starts > .5 + eps)
  set.seed(i)
  manifold.clustering(Xhat, 2, 
                      parallel = TRUE, 
                      intercept = intercept, 
                      maxit = maxit,
                      normalize = normalize, 
                      eps = 1e-3,
                      verbose = FALSE,
                      animate = animate,
                      animation.dir = animation.dir,
                      animation.title = paste('random', i, ifelse(intercept,
                                                                  'w-intercept',
                                                                  'no-intercept'),
                                              sep = '-'))
})
names(out) <- paste('random', seq(random.starts))
names(out) <- sapply(seq(random.starts), function(i) {
  if (i / random.starts < .5 + eps) {
    paste('random', i, '\n(w/o intercept)')
  } else {
    paste('random', i, '\n(w/ intercept)')
  }
})

out[['spectral clust\n(w/o intercept)']] <- manifold.clustering(
  Xhat, 2, 
  parallel = TRUE, 
  intercept = FALSE, 
  normalize = normalize, 
  maxit = maxit, 
  eps = 1e-3,
  initialization = 'spectral',
  verbose = FALSE,
  animate = animate,
  animation.dir = animation.dir,
  animation.title = 'spec-clust-no-intercept')

out[['spectral clust\n(w/ intercept)']] <- manifold.clustering(
  Xhat, 2,
  parallel = TRUE,
  intercept = TRUE,
  normalize = normalize, 
  maxit = maxit,
  eps = 1e-3,
  initialization = 'spectral',
  verbose = FALSE,
  animate = animate,
  animation.dir = animation.dir,
  animation.title = 'spec-clust-w-intercept')

out[['ground truth\n(w/o intercept)']] <- manifold.clustering(
  Xhat, 2,
  parallel = TRUE,
  intercept = FALSE,
  normalize = normalize, 
  maxit = maxit,
  eps = 1e-3,
  initialization = z,
  verbose = FALSE,
  animate = animate,
  animation.dir = animation.dir,
  animation.title = 'groundtruth-no-intercept')

@@

```{r cache = TRUE}
set.seed(123)
z.spectral <- spectral.clustering(A)
macaque.out <- manifold.clustering(Xhat, 2, 
                                   # initialization = 'random',
                                   initialization = z.spectral,
                                   # initialization = z,
                                   intercept = FALSE, 
                                   parallel = TRUE, 
                                   verbose = FALSE, 
                                   animate = animate, 
                                   curve.init = 'x',
                                   animation.title = 'macaque')
```

@@

```{r}
plot(macaque.out$loss, type = 'l')
```

@@

```{r cache = TRUE}
z.init <- hclust(dist(Xhat), method = 'single') %>% 
  cutree(2)
nonintersect.out <- manifold.clustering(Xhat, 2, 
                                        # initialization = z.init,
                                        initialization = 'random', 
                                        intercept = TRUE, 
                                        animate = animate,
                                        animation.title = 'nonintersect')
```

@@

ggplot() + 
  geom_point(aes(x = Xhat[, 1], y = Xhat[, 2], 
                 colour = factor(nonintersect.out$z)),
             size = .5) + 
  labs(x = expression(x[1]), y = expression(x[2]), colour = NULL) + 
  theme_bw() + 
  coord_fixed() +
  theme(legend.position = 'none') +
  geom_path(aes(x = Y1[, 1], y = Y1[, 2])) +
  geom_path(aes(x = Y2[, 1], y = Y2[, 2]))
```
